NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models: Supplemental Materials

### 2018-06-22 paper/attention.tex
remove related works

As illustrated in Fig.~\ref{fig:attentionVis}, we employ two visual encodings for visualizing attention. In the graph attention view (Fig.~\ref{fig:attentionVis}(a)), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. The color of the rectangle text block encodes the sum of all edge values connected to it (darker shade of blues correspond to higher values).
%
The graph view is good for highlighting the most dominant alignments. However, if many attention values are high, the edges may become cluttered, leading to less effective visualization. Also, if the sentence structures between premise and hypothesis are drastically different, we are likely to see the prominent edges cross each other, which can also lead to confusing and misleading visual patterns.

The matrix attention view (Fig.~\ref{fig:attentionVis}(b)), despite being more verbose and less efficient in highlighting the dominant alignment, does not have similar shortcomings. Together, the graph and matrix views complement each other and provide the same information from different perspectives. As illustrated by the pink arrowed lines, we can see how the same attention value is visualized in both the graph and the matrix view.
To help the user recognize the correspondence during the exploration, we enable the linkage between highlighted actions in both views (see Fig.~\ref{fig:attentionVis}(a)(b), the attention of the two ``couple'' is highlighted).

@@

Even though the attention does not explicitly encode any grammar, it often highlights essential words in the sentence structure.
%
To help the researcher better understand the relationship between attention and sentence structure, as illustrated in Fig.~\ref{fig:attentionVis}($a_{1}$), we overlay the grammar dependency tree~\cite{Nivre2005} structure next to the sentence.
%
Since the dependency tree encodes the word importance information in a hierarchical matter, it is very suitable for sentence simplification tasks.
Therefore, we utilize the grammar dependency tree to simplify the attention representations to combat the visual clutter when examining long sentences (see Fig.~\ref{fig:attentionVis}(d)).
A simplification example is shown in  Fig.~\ref{fig:attentionVis}(d)(e)(f).

>>>>

As illustrated in Fig.~\ref{fig:attentionVis}, we employ two visual encodings for visualizing attention. In the graph attention view (Fig.~\ref{fig:attentionVis}(a)), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. The color of the rectangle text block encodes the sum of all edge values connected to it (darker shade of blues correspond to higher values).
%
The graph view is good for highlighting the most dominant alignments. However, if many attention values are high, the edges may become cluttered, leading to less effective visualization. Also, if the sentence structures between premise and hypothesis are drastically different, we are likely to see the prominent edges cross each other, which can also lead to confusing and misleading visual patterns.
%
The matrix attention view (Fig.~\ref{fig:attentionVis}(b)), despite being more verbose and less efficient in highlighting the dominant alignment, does not have similar shortcomings. However, extra effort may be required for identifying the words corresponds to high-value entries in the matrix. Together, the graph and matrix views complement each other and provide the same information from different perspectives (a related matrix/graph visualization scheme is presented in \cite{MaKenyonForbes2015} for studying brain network). 
%As illustrated by the pink arrowed lines, we can see how the same attention value is visualized in both the graph and the matrix view.
To help the user recognize the correspondence during the exploration, we enable the linkage between highlighted actions in both views (see Fig.~\ref{fig:attentionVis}(a)(b), the attention of the two ``couple'' is highlighted).

@@

Even though the attention does not explicitly encode any grammar, it often highlights essential words in the sentence structure.
%
To help the researcher better understand the relationship between attention and sentence structure, as illustrated in Fig.~\ref{fig:attentionVis}($a_{1}$), we overlay the grammar dependency tree~\cite{Nivre2005} structure next to the sentence.
%
Since the dependency tree encodes the word importance information in a hierarchical manner, it is very suitable for sentence simplification tasks.
Therefore, we utilize the grammar dependency tree to simplify the attention representations to combat the visual clutter when examining long sentences (see Fig.~\ref{fig:attentionVis}(d)). A simplification example is shown in  Fig.~\ref{fig:attentionVis}(d)(e)(f).

### 2018-06-22 paper/intro.tex
remove related works

Compared to image classification tasks, natural language processing (NLP) systems often involve additional challenges such as the discrete nature of words.
For example, \emph{feature visualization}~\cite{OlahMordvintsevSchubert2017}, an often deployed technique for illustrating what image features a given part of the network (e.g., neuron, layer, or channel) captures, cannot be readily generalized to natural language.
\shusen{
An image (pixel values) corresponds to a continuous solution space. which is more accessible for optimization and human visual recognition (i.e., the existence or the absence of patterns).
For natural language, even though we can encode a word as a vector~\cite{MikolovSutskeverChen2013, PenningtonSocherManning2014}, the word embedding space is still discretely defined, in which the interpolation of two vectors (i.e., two words) does not hold clear meaning.}
%
These restrictions call for new avenues for solving the interpretation challenges of natural language models, which motivates the proposed work.

>>>>

Compared to image classification tasks, natural language processing (NLP) systems often involve additional challenges such as the discrete nature of words.
For example, \emph{feature visualization}~\cite{OlahMordvintsevSchubert2017}, an often deployed technique for illustrating what image features a given part of the network (e.g., neuron, layer, or channel) captures, cannot be readily generalized to natural language.
%\shusen{
An image (pixel values) corresponds to a continuous solution space. which is more accessible for optimization and human visual recognition (i.e., the existence or the absence of patterns).
For natural language, even though we can encode a word as a vector~\cite{MikolovSutskeverChen2013, PenningtonSocherManning2014}, the word embedding space is still discretely defined, in which the interpolation of two vectors (i.e., two words) does not hold clear meaning.%}
%
These restrictions call for new avenues for solving the interpretation challenges of natural language models, which motivates the proposed work.

### 2018-06-22 paper/relatedWorks.tex
remove related works

\textbf{Model Interpretability.}
Before one can solve the model interpretability challenge, it worth to understand what constitutes an explanation for a machine learning model.
As discussed in~\cite{Lipton2016, Doshi-Velez2017}, there is little consensus on what interpretability is in the context of machine learning. A taxonomy is suggested~\cite{Doshi-Velez2017} to promote a more rigorous study on such a subject. In the work by Lipton~\cite{Lipton2016}, many philosophical questions related to model interpretability are examined and the study also challenged the commonly held belief that linear models are more interpretable than deep neural networks.
%
In the machine learning community, one of the most studied areas in model interpretability is the interpretation of the convolution neural network (CNN) ~\cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017, OlahSatyanarayanJohnson2018}.
%
These methods either identify the pattern that maximally activates a part of the network (referred to as \emph{feature visualization}, i.e., what feature does the given neuron, channel, or layer captures?), or reveals what areas in the input that contribute most to the predicted label (often referred to as \emph{attribution}, i.e., attribute the prediction label in the input domain).
%
Besides the techniques designed for specific types of models, a few work~\cite{RibeiroSinghGuestrin2016, KrausePererNg2016, LundbergLee2017} approach the interpretability challenge from a model agnostic perspective, which makes them applicable to different applications. In the LIME~\cite{RibeiroSinghGuestrin2016} work, a simpler linear model is fitted near the prediction of interests to provide a simpler but easier to understand snapshot of the classifier near a specific input.
%
Despite being invaluable for generating a human understandable explanation, however, the lack of the ability to access and explore the internal states of the system limits their application.
%which is vitally important for making sense of the inner works of a model,
To avoid similar limitations, the proposed system enables direct access and manipulation of the internals of the model.

>>>>

\textbf{Model Interpretability.}
%Before one can solve the model interpretability challenge, it worth to understand what constitutes an explanation for a machine learning model.
%As discussed in~\cite{Lipton2016, Doshi-Velez2017}, there is little consensus on what interpretability is in the context of machine learning. A taxonomy is suggested~\cite{Doshi-Velez2017} to promote a more rigorous study on such a subject. In the work by Lipton~\cite{Lipton2016}, many philosophical questions related to model interpretability are examined and the study also challenged the commonly held belief that linear models are more interpretable than deep neural networks.
%
In the machine learning community, one of the most studied areas in model interpretability is the interpretation of the convolution neural network (CNN) ~\cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017, OlahSatyanarayanJohnson2018}.
%
These methods either identify the pattern that maximally activates a part of the network (referred to as \emph{feature visualization}, i.e., what feature does the given neuron, channel, or layer captures?), or reveals what areas in the input that contribute most to the predicted label (often referred to as \emph{attribution}, i.e., attribute the prediction label in the input domain).
%
Besides the techniques designed for specific types of models, a few work~\cite{RibeiroSinghGuestrin2016, KrausePererNg2016, LundbergLee2017} approach the interpretability challenge from a model agnostic perspective, which makes them applicable to different applications. In the LIME~\cite{RibeiroSinghGuestrin2016} work, a simpler linear model is fitted near the prediction of interests to provide a simpler but easier to understand snapshot of the classifier near a specific input.
%
Despite being invaluable for generating a human understandable explanation, however, the lack of the ability to access and explore the internal states of the system limits their application.
%which is vitally important for making sense of the inner works of a model,
To avoid similar limitations, the proposed system enables direct access and manipulation of the internals of the model.

### 2018-06-22 paper/task.tex
remove related works

\section{Task Analysis}
\label{sec:task}
Identifying the analysis goals and the corresponding visualization tasks is the first significant step in designing a useful tool. In this section, we examine how domain experts normally approach the task of analyzing a model (i.e., assess its strength and weakness) and identify where and how an interactive visualization system can aid in such a process.
Then, we distill the numerous discussions with domain experts into concrete tasks, which guide the design and implementation of the proposed tool.

During our long-term collaboration with NLP experts, we have conducted extensive discussions to understand the common approach employed by researchers for assessing the behavior of a model.%
The predictive accuracy has its place as an objective evaluation metric to measure the overall effectiveness of the model; however, the accuracy number alone does not provide the full story.
For example, the model may produce correct predictions for the ``wrong'' reasons (e.g., pick up an unintended pattern in the training data that cannot be generalized in real-word scenarios).
%
As a result, the domain experts rely on an exploration-centric approach to conduct error analysis and obtain intuition.

There is little one can learn when everything is behaving as expected, and therefore,
observing how models make mistakes under various scenarios is at the center of such a process.
The experts often start with simple failure examples, and then make minor perturbations (replace a word or phrase) to the input and observe the change in the prediction. This exercise helps isolate where in the input likely contribute to the failures. For many NLP models, the attention information (see Section~\ref{sec:attention}) is essential to infer the mechanism of the model. Experts often print out the attention values or generate plots to visualize sentence alignment or compare attention values. As the experts explore more variation of similar examples, combined with their domain knowledge, they may develop hypotheses about the cause of failures, or ask additional questions that lead to further experimentation.

In such an exploration workflow, NLP researchers often need to utilize multiple scripts and manually run all the experiments. The batch process based approach, is not only time consuming, but also hinders the flow of reasoning which rely on obtaining instantaneous feedbacks, and making many on-the-fly adjustments.
%
Therefore, the expert driven exploration process can greatly benefit from the introduction of an interactive visual exploration environment, where the researchers can easily express their hypotheses and obtain quick visual feedback of the results.
%
Moreover, by introducing new visual encodings and summarizations, we can drastically expand the ways domain experts interact with the model, enable exploration options that previously were not possible either due to tedious manual operation or the lack of communication channels.

%Through many rounds of discussions with NLP experts,
%In such a highly exploratory process, the expert likely will have many ``what if'' type of questions (e.g., what if I make change to this word in the input, what if I force the words to align correctly in the attention), get responsive answers to these queries is essential to maintain the flow of reasoning.
%
To support the exploratory workflow for analyzing NLI models, we need to address the following tasks.

\begin{itemize}

\item \textbf{T1}: Understand the stability of a prediction, i.e., how perturbations of the input affect the behavior of the model?

@@

\item \textbf{T3}: When the predicted label is wrong, how can we update the model to correct the prediction? What are the effects of updating different parts of the model?

>>>>

\section{Task Analysis}
\label{sec:task}
%Identifying the analysis goals and the corresponding visualization tasks is the first significant step in designing a useful tool. 
%In this section, we examine how domain experts normally approach the task of analyzing a model (i.e., assess its strength and weakness) and identify where and how an interactive visualization system can aid in such a process.
%\shusen{
The primary driving force for designing the proposed tool is the error analysis challenges faced by our long-term collaborators working on natural language inference research. During the entire design and development process, we worked closely with two NLP experts via weekly meetings in a period of roughly seven month, their constant evaluation and feedback have helped shape the tool we see today.
%}
%Then, we distill the numerous discussions with domain experts into concrete tasks, which guide the design and implementation of the proposed tool.
%During our long-term collaboration with two NLP experts, 
During this period, we have conducted extensive discussions to understand the common approaches employed by researchers for assessing the behavior of a model.
% as well as provide feedback on the various prototyping stages of the proposed visualization.%
The predictive accuracy has its place as an objective evaluation metric to measure the overall effectiveness of the model; however, the accuracy number alone does not provide the full story.
For example, the model may produce correct predictions for the ``wrong'' reasons (e.g., pick up an unintended pattern in the training data that cannot be generalized in real-word scenarios).
%
As a result, the domain experts rely on an exploration-centric approach to conduct error analysis and obtain intuition.
%
There is little one can learn when everything is behaving as expected, and therefore,
observing how models make mistakes under various scenarios is at the center of such a process.
The experts often start with simple failure examples, and then make minor perturbations (replace a word or phrase) to the input and observe the change in the prediction. This exercise helps isolate where in the input likely contribute to the failures. For many NLP models, the attention information (see Section~\ref{sec:attention}) is essential to infer the mechanism of the model. Experts often print out the attention values or generate plots to visualize sentence alignment or compare attention values. As the experts explore more variation of similar examples, combined with their domain knowledge, they may develop hypotheses about the cause of failures, or ask additional questions that lead to further experimentation.

In such an exploration workflow, NLP researchers often need to utilize multiple scripts and manually run all the experiments. The batch process based approach, is not only time consuming, but also hinders the flow of reasoning which rely on obtaining instantaneous feedbacks, and making many on-the-fly adjustments.
%
Therefore, the expert driven exploration process can greatly benefit from the introduction of an interactive visual exploration environment, where the researchers can easily express their hypotheses and obtain quick visual feedback of the results.
%
Moreover, by introducing new visual encodings and summarizations, we can drastically expand the ways domain experts interact with the model, enable exploration options that previously were not possible either due to tedious manual operation or the lack of communication channels.
%
%In such a highly exploratory process, the expert likely will have many ``what if'' type of questions (e.g., what if I make change to this word in the input, what if I force the words to align correctly in the attention), get responsive answers to these queries is essential to maintain the flow of reasoning.
%\shusen{
To support the exploratory workflow for analyzing NLI models,
we aim to address the following tasks based on our extensive discussions with NLP experts.
%}

\begin{itemize}
    \item \textbf{T1}: Understand the stability of a prediction, i.e., how perturbations of the input affect the behavior of the model?

@@

    \item \textbf{T3}: When the predicted label is wrong, how can we update the model to correct the prediction? What are the effects of updating different parts of the model?
\end{itemize}

In Section~\ref{sec:caseStudy}, we will discussion in detail how we can utilize the proposed tool to address these tasks (the application scenarios 1,2,3 corresponds to \textbf{T1-T3}, respectively).
